---
title: "Local field enhancement factors for a dipole near or inside a multilayer"
date: "`r format(Sys.time(), '%d %B, %Y')`"
author: "baptiste Auguié"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    fig_width: 7
    fig_height: 4
    fig_caption: true
vignette: >
  %\VignetteIndexEntry{field_multilayer}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r demo, message=FALSE, echo=FALSE}
knitr::read_demo("field_multilayer", package="planar")
```
```{r load, echo=FALSE,results='hide'}
```

#### Setting up

```{r setup, results='hide'}
```

We simulate a thin metal layer sandwiched between glass (incident medium) and air, illuminated from the glass side under p-polarisation, at 44 degrees (Kretschmann configuration). The electric field is normalised by the permittivity of each layer to exemplify the continuity of electric displacement (squared, here) across interfaces.

```{r simulation}
```
A strong electric field can be noted, particularly for the perpendicular component, which decays exponentially from the metal-air interface: this is characteristic of the excitation of surface plasmon-polaritons.

```{r plot, fig.width=12}
```
