---
title: "Basic usage of the tamm package"
output:
  md_document:
    fig_width: 7
    fig_height: 3
    toc: true
    toc_depth: 2
    variant: "markdown_github"
author: baptiste Auguié
---

```{r demo, message=FALSE, echo=FALSE}
library(knitr)
library(ggplot2)
library(reshape2)
library(dielectric)
opts_chunk$set(fig.path='')
knitr::read_chunk("basic.r")
```
We start by loading the `tamm` package, 
```{r setup, message=FALSE, echo=TRUE}
```
and creating a multilayer structure via the high-level `tamm_stack` function (`tamm_stack_ir` has suitable default values for typical AlAs/GaAs semi-conductors used in the near-infrared). We can straight-forwardly visualise the structure by plotting it,
```{r structure, message=FALSE, echo=TRUE}
```
The far-field properties (reflectance, transmittance, absorbance) can be simulated with the following function, the output in long-format being directly fed to ggplot2,
```{r ff, message=FALSE, echo=TRUE}
```
The spectral position of the Tamm mode having been isolated, we can visualise the electric field profile inside the structure and superpose it to the geometry,
```{r nf, message=FALSE, echo=TRUE}
```
