Reproducing Fig. 6.6, p. 188 of Mac Leod's Thin Film Optical Filters.
The structure is a stack of lambda/4 layers of alternating indices nH (high) and nL (low) on a glass substrate. 

```{r , echo=FALSE,results='hide'}
opts_chunk$set(fig.path="bragg/",
               warning=FALSE,error=FALSE,message=FALSE,tidy=TRUE)
library(ggplot2)
theme_set(theme_minimal() + theme(panel.border=element_rect(fill=NA)))
```

#### Setting up

```{r , results='hide'}
library(planar)
library(ggplot2)
require(reshape2)
require(plyr)

makeStack <- function(n = 3, lambda=seq(200, 1000),
                      lambda0 = 460, thickness = lambda0/4, 
                      nH=2.3, nL=1.38, nS=1.52,
                      angle=0){

  epsilon.list <- c(1, rep(c(nL, nH)^2, n), nS)
  thickness.list <- c(0, rep(thickness/c(nL, nH), n), 0)
  
  params <- list(epsilon=as.list(epsilon.list),
                 lambda=lambda, thickness=thickness.list,
                 theta=angle*pi/180, polarisation='p')
  
  data.frame(do.call(recursive_fresnelcpp, params))
}
```
With increasing number of layers, the reflectivity stop-band becomes stronger, with a more rectangular transfer function. Side-bands also develop accordingly, however, with higher and higher frequency.

```{r simulation}

params <- expand.grid(n=seq(5,20,by=2), angle=seq(0,60, by=30))
all <- mdply(params, makeStack)

ggplot(all)+ facet_grid(angle~.)+
  geom_line(aes(2*pi/k0, R, colour=n, group=n))+
  labs(colour="layers") +
  scale_x_continuous("wavelength /nm")+
  scale_y_continuous("Reflectivity", expand=c(0,0), lim=c(0,1))

```
