Local field enhancement factors for a dipole near or inside a multilayer.

```{r , echo=FALSE,results='hide'}
opts_chunk$set(fig.path="internalfield/",
               warning=FALSE,error=FALSE,message=FALSE,tidy=FALSE)
library(ggplot2)
theme_set(theme_minimal() + theme(panel.border=element_rect(fill=NA)))
```

#### Setting up

```{r , results='hide'}
library(planar)
library(ggplot2)
require(reshape2)
require(plyr)
```
We simulate a thin metal layer sandwiched between glass (incident medium) and air, illuminated from the glass side under p-polarisation, at 44 degrees (Kretschmann configuration). The electric field is normalised by the permittivity of each layer to exemplify the continuity of electric displacement (squared, here) across interfaces.

```{r simulation}
m <- field_profile(lambda=633, theta=44*pi/180, polarisation='p',
                   thickness=c(0, 50, 0), dmax=400, res=5000,
                   epsilon=list(1.5^2, -12+1i, 1.0^2), 
                   displacement=TRUE)
m$L1 <- factor(m$L1, labels=c("glass", "metal", "air"))
str(m)
```
A strong electric field can be noted, particularly for the perpendicular component, which decays exponentially from the metal-air interface: this is characteristic of the excitation of surface plasmon-polaritons.
```{r plot, fig.width=12}

limits <- ddply(m, .(L1), summarize, 
                xmin=min(x), xmax=max(x), ymin=-Inf, ymax=Inf)

  ggplot(m) +
  facet_grid(variable~., scales="free")+
  geom_rect(aes(xmin=xmin, ymin=ymin, xmax=xmax, ymax=ymax, 
                fill=factor(L1)),     data=limits, alpha=0.2) +
  geom_path(aes(x, value, colour=factor(L1))) +
  scale_x_continuous("x /nm",expand=c(0,0)) +
  scale_y_continuous("LFIEF",expand=c(0,0)) +
  geom_hline(yintercept=0) +
  scale_colour_brewer("", palette="Set1")+
  scale_fill_brewer("", palette="Pastel1") +
  theme_minimal()

```
