## Near-field profile under gaussian-beam illumination
_baptiste Auguié -- `r format(Sys.time(), "%d %B, %Y")`_

Using the angular spectrum decomposition of a gaussian beam in a bundle of plane waves, the near-field profile is calculated on the transmitted side of an interface illuminated under total internal reflection.

```{r , echo=FALSE,results='hide'}
library(knitr)
library(ggplot2)
library(rgl)
knit_hooks$set(rgl = function(before, options, envir) {
  # if a device was opened before this chunk, close it
  if (before && rgl.cur() > 0) rgl.close()
  hook_rgl(before, options, envir)
})
opts_chunk$set(fig.path="gaussianbeam/", cache=TRUE, cache.path="gaussian/",
               warning=FALSE,error=FALSE,message=FALSE,tidy=FALSE)
library(ggplot2)
theme_set(theme_minimal() + theme(panel.border=element_rect(fill=NA)))
```

#### Setting up

```{r , results='hide'}
library(planar)
library(ggplot2)
library(plyr)
w0 <- 1e3 # beam waist 1 micron
alpha <- 70*pi/180 # incident angle
```

## Depth profile

```{r depth}
xyz <- as.matrix(expand.grid(x=seq(0,1e4,length=20), y=0, z=seq(1, 500, length=50)))
res <- adply(xyz, 1, gaussian_field, d=50, nl=sqrt(epsAu(633)$epsilon),
             wavelength=633, w0=5e3, ni=1.5+0i, alpha=45, maxEval=1000)
xyz <- data.frame(xyz, field=res[[2]])
ggplot(xyz, aes(z, field, colour=x, group=x))+
  geom_line() +scale_y_log10()+
  labs(x=expression("z /nm"), y=expression(log("|E|"^2)), 
       colour=expression("x /nm")) 

```
We observe that the field decays exponentially away from the interface with the same characteristic distance across the beam profile.

## Goos-Hanshen shift

Under total-internal reflection, the beam undergoes a spatial displacement known as Goos-Hanshen shift. The centre of the beam is displaced from the origin.

```{r GH}
w0 <- 1000
xyz <- as.matrix(expand.grid(x=seq(-5*w0, 15*w0,length=100), y=0, z=seq(0, 500, by=50)))
res <- adply(xyz, 1, gaussian_field, d=50, nl=sqrt(epsAu(633)$epsilon),
             wavelength=633, w0=w0, ni=1.5+0i, alpha=44.5*pi/180, maxEval=2000)
xyz <- data.frame(xyz, field=res[[2]])

ggplot(xyz, aes(x, field, colour=z, group=z))+
  geom_line() +
  geom_vline(aes(x=0,y=NULL),lty=2) +
  labs(x=expression("x /nm"), y=expression("|E|"^2), 
       colour=expression("z /nm")) 
```

## 2D map

We can calculate the field at a grid of points in 2D or 3D. Here is the result for a slice at $z=10$nm.

```{r map}
w0 <- 2000
xyz <- as.matrix(expand.grid(x=seq(-3*w0, 5*w0, length=50), 
                             y=seq(-3*w0, 3*w0, length=50),
                             z=10))
res <- adply(xyz, 1, gaussian_field, d=50, nl=sqrt(epsAu(633)$epsilon),
             wavelength=633, w0=w0, ni=1.5+0i, alpha=44.5*pi/180, maxEval=500)
xyz <- data.frame(xyz, field=res[[2]])

  ggplot(xyz, aes(x/1e3, y/1e3, fill=field))+
  geom_raster(interpolate=TRUE) +
  scale_x_continuous(expand=c(0,0))+
  scale_y_continuous(expand=c(0,0)) +
  labs(x=expression("x /nm"), fill=expression("|E|"^2), 
       y=expression("y /nm")) +
  coord_fixed()

```

```{r volume, rgl=TRUE}
w0 <- 2000
x <- seq(-5*w0, 15*w0, length=100)
z <- seq(2, 1000, length=100)
xyz <- as.matrix(expand.grid(x=x, 
                             y=0, #seq(-3*w0, 3*w0, length=30),
                             z=z))
                             
res <- adply(xyz, 1, gaussian_field, d=50, nl=sqrt(epsAu(633)$epsilon),
             wavelength=633, w0=w0, ni=1.5+0i, alpha=44.5*pi/180, maxEval=500)
xyz <- data.frame(xyz, field=res[[2]])

col <- heat.colors(100)[cut(xyz$field, 100)]

rgl.surface(x=x, y=matrix(xyz$field/max(xyz$field)*1e4, ncol=100), 
            z=z*10, color=col, back="lines", lit=FALSE, alpha=0.9)
# 
# sprites3d(xyz[,1], xyz[,2], xyz[,3]*10 ,radius=1000,lit=FALSE, alpha=.1,
#         color=grey(scales::rescale(xyz[,4], to=c(0.9,0))))
```

The effect of the beam size on the field enhancement is quite dramatic as soon as the spread of incident angles causes the coupling to SPPs to deteriorate. Here we simulate the LFIEF for silver, with a beam waist varying from 1000 microns to 1 micron.

```{r waist, fig.height=15}
## first, check the plane wave result
results <- multilayer(epsilon=list(1.52^2, epsAg(632.8)$epsilon, 1.0),
                         wavelength=632.8, thickness=c(0, 50, 0), d=1,
                         angle=seq(0, pi/2, length=2e3), polarisation='p')

max(results$Mr.perp[[2]], na.rm=T)

spp <- results$angle[which.min(results$R)]

simul <- function(w0=10){
  w0 <- w0*1e3
  xyz <- as.matrix(expand.grid(x=seq(-5*w0, 5*w0+5000,length=100), y=0, z=seq(0, 300, by=50)))
  res <- adply(xyz, 1, gaussian_field, d=50, nl=sqrt(epsAg(632.8)$epsilon),
               wavelength=632.8, w0=w0, ni=1.52, alpha=spp, maxEval=1000)
  data.frame(xyz, field=res[[2]])
}
# 

all <- mdply(data.frame(w0=c(1e3, 5e2, 1e2, 50, 10, 5, 1)), simul)

ggplot(all, aes(x/w0/1000, field, group=z, colour=z))+
  geom_line() + facet_grid(w0~., scales="free", labeller=label_both) +
  geom_vline(aes(x=0,y=NULL),lty=2) +
  labs(x=expression(x/w[0]), y=expression("|E|"^2), 
       colour=expression("z /nm")) +
  coord_cartesian(xlim=c(-5,5)) + theme_minimal()+
  guides(colour=guide_legend())
```


